---
title: "Quiz 3"
author: "S. Servaes"
date: "3/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
library(ggplot2)
```

## Question 1

```{r question 1}
data(mtcars)
fit <- lm(mpg ~ factor(cyl) + wt, mtcars)
summary(fit)$coefficients
```

Answer is -6.0709.

## Question 2

```{r question 2}
data(mtcars)
fit1 <- lm(mpg ~ factor(cyl), mtcars)
fit2 <- lm(mpg ~ factor(cyl) + wt, mtcars)

summary(fit1)$coefficients
summary(fit2)$coefficients
```

Holding weight constant, cylinders appear to have less of an impact on mpg than if weight is disregarded.

## Question 3

```{r question 3}
data(mtcars)
fit1 <- lm(mpg ~ factor(cyl) + wt, mtcars)
fit2 <- lm(mpg ~ factor(cyl) * wt, mtcars)
anova(fit1, fit2)
```

The P-value is larger than 0.05. So, according to our criterion, we would fail to reject, which suggests that the interaction terms may not be necessary.

## Question 4

The estimated expected change in MPG per one ton increase in weight for for a specific number of cylinders (4, 6, 8).

## Question 5

```{r question 5}
x <- c(0.586, 0.166, -0.042, -0.614, 11.72)
y <- c(0.549, -0.026, -0.127, -0.751, 1.344)

fit <- lm(y ~x)
hatvalues(fit)
plot(fit, 4)
```

Answer is 0.9946.

## Question 5

```{r question 5}
x <- c(0.586, 0.166, -0.042, -0.614, 11.72)
y <- c(0.549, -0.026, -0.127, -0.751, 1.344)

fit <- lm(y ~x)
dfbetas(fit)
```

Answer is -133.82.